---
title: "Bring Out Your Dead"
author: "Yva Eline"
date: "10/22/2024"
format: 
  html:
    fig-width: 8
    fig-height: 4
    code-fold: true
---

# Summary of Experiment

Larvae were subjected to a series of treatments with "Lactobacillus spp" and monitored for viability over a six days. Survivability grafts were generated to assess the consistency of the experiment.

# Packages and Libraries

Look, man, you gotta install the packages and libraries somewhere... AND HERE THEY ARE!

```{r libraries, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

install.packages("readxl") install.packages("dplyr") install.packages("ggplot2")

library(readxl) library(dplyr) library(ggplot2)

# Conclusion

# read data from each sheet & add a replicate column

rep1 \<- read_excel("larvae_ncimb_lacto.xlsx", sheet = "replicate_1") %\>% mutate(replicate = "Replicate 1") rep2 \<- read_excel("larvae_ncimb_lacto.xlsx", sheet = "replicate_2") %\>% mutate(replicate = "Replicate 2") rep3 \<- read_excel("larvae_ncimb_lacto.xlsx", sheet = "replicate_3") %\>% mutate(replicate = "Replicate 3") rep4 \<- read_excel("larvae_ncimb_lacto.xlsx", sheet = "replicate_4") %\>% mutate(replicate = "Replicate 4")

# combine data into one dataframe

combined_data \<- bind_rows(rep1, rep2, rep3, rep4)

# convert treatment labels to upper case - because I am a dummy

combined_data \<- combined_data %\>% mutate(treatment = toupper(treatment))

# inspect

print(head(combined_data)) print(str(combined_data))

# summarize data: calculate mean and standard error for alive larvae by treatment and day

summary_data \<- combined_data %\>% group_by(treatment, day) %\>% summarize( mean_alive_count = mean(alive_count, na.rm = TRUE), se_alive_count = sd(alive_count, na.rm = TRUE) / sqrt(n()) )

#inspect summarized data - because here thar might be fuck ups, yarrr print(summary_data)

# combined plot with error bars

combined_plot \<- ggplot(summary_data, aes(x = day, y = mean_alive_count, color = treatment, group = treatment)) + geom_line() + \# Line graph geom_point() + \# Points on the line graph geom_errorbar(aes(ymin = mean_alive_count - se_alive_count, ymax = mean_alive_count + se_alive_count), width = 0.2) + \# Error bars labs(title = "Number of Alive Larvae Over Time by Treatment with Error Bars", x = "Day", y = "Mean Number of Alive Larvae") + scale_color_manual( values = c("1000" = "red", "100" = "blue", "10" = "green", "NT" = "black") ) + theme_minimal() + theme( legend.title = element_blank(), legend.text = element_text(size = 10), plot.background = element_rect(fill = "white", color = NA), \# Ensure white background panel.background = element_rect(fill = "white", color = NA) \# Ensure white background )

# view and save combined plot as .png

print(combined_plot) ggsave("combined_plot.png", plot = combined_plot, width = 8, height = 6, bg = "white")

# now for the individual replicates

replicates \<- unique(combined_data\$replicate)

for (rep in replicates) { replicate_data \<- combined_data %\>% filter(replicate == rep)

p \<- ggplot(replicate_data, aes(x = day, y = alive_count, color = treatment, group = treatment)) + geom_line() + \# Line graph geom_point() + \# Points on the line graph labs(title = paste("Number of Alive Larvae Over Time (", rep, ")", sep = ""), x = "Day", y = "Number of Alive Larvae") + scale_color_manual( values = c("1000" = "red", "100" = "blue", "10" = "green", "NT" = "black") ) + theme_minimal() + theme( legend.title = element_blank(), legend.text = element_text(size = 10), plot.background = element_rect(fill = "white", color = NA), \# Ensure white background panel.background = element_rect(fill = "white", color = NA) \# Ensure white background )

\# view and save replicates as .png print(p) ggsave(paste0("plot\_", gsub(" ", "\_", rep), ".png"), plot = p, width = 8, height = 6, bg = "white") }
