---
title: "testing_my_patience"
author: "Yva"
date: "2024-10-22"
format: "html"
html:
  fig_width: 8
  fig-height: 4
---

# Summary

# Packages & Libraries Area

```{r}

```

## Load in that sweet sweet spreadsheet and marinate it a bit so that it reads easier


```{r cars}


```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.


install.packages("readxl")
install.packages("dplyr")
install.packages("ggplot2")

library(readxl)
library(dplyr)
library(ggplot2)

# read data from each sheet & add a replicate column
rep1 <- read_excel("larvae_ncimb_lacto.xlsx", sheet = "replicate_1") %>% mutate(replicate = "Replicate 1")
rep2 <- read_excel("larvae_ncimb_lacto.xlsx", sheet = "replicate_2") %>% mutate(replicate = "Replicate 2")
rep3 <- read_excel("larvae_ncimb_lacto.xlsx", sheet = "replicate_3") %>% mutate(replicate = "Replicate 3")
rep4 <- read_excel("larvae_ncimb_lacto.xlsx", sheet = "replicate_4") %>% mutate(replicate = "Replicate 4")

# combine data into one dataframe
combined_data <- bind_rows(rep1, rep2, rep3, rep4)

# convert treatment labels to upper case - because I am a dummy
combined_data <- combined_data %>% mutate(treatment = toupper(treatment))

# inspect
print(head(combined_data))
print(str(combined_data))

# summarize data: calculate mean and standard error for alive larvae by treatment and day
summary_data <- combined_data %>%
  group_by(treatment, day) %>%
  summarize(
    mean_alive_count = mean(alive_count, na.rm = TRUE),
    se_alive_count = sd(alive_count, na.rm = TRUE) / sqrt(n())
  )

#inspect summarized data - because here thar might be fuck ups, yarrr
print(summary_data)

# combined plot with error bars
combined_plot <- ggplot(summary_data, aes(x = day, y = mean_alive_count, color = treatment, group = treatment)) +
  geom_line() +  # Line graph
  geom_point() + # Points on the line graph
  geom_errorbar(aes(ymin = mean_alive_count - se_alive_count, ymax = mean_alive_count + se_alive_count), width = 0.2) + # Error bars
  labs(title = "Number of Alive Larvae Over Time by Treatment with Error Bars",
       x = "Day",
       y = "Mean Number of Alive Larvae") +
  scale_color_manual(
    values = c("1000" = "red", "100" = "blue", "10" = "green", "NT" = "black")
  ) +
  theme_minimal() +
  theme(
    legend.title = element_blank(), 
    legend.text = element_text(size = 10),
    plot.background = element_rect(fill = "white", color = NA),  # Ensure white background
    panel.background = element_rect(fill = "white", color = NA)  # Ensure white background
  )

# view and save combined plot as .png
print(combined_plot)
ggsave("combined_plot.png", plot = combined_plot, width = 8, height = 6, bg = "white")

# now for the individual replicates
replicates <- unique(combined_data$replicate)

for (rep in replicates) {
  replicate_data <- combined_data %>% filter(replicate == rep)
  
  p <- ggplot(replicate_data, aes(x = day, y = alive_count, color = treatment, group = treatment)) +
    geom_line() +  # Line graph
    geom_point() + # Points on the line graph
    labs(title = paste("Number of Alive Larvae Over Time (", rep, ")", sep = ""),
         x = "Day",
         y = "Number of Alive Larvae") +
    scale_color_manual(
      values = c("1000" = "red", "100" = "blue", "10" = "green", "NT" = "black")
    ) +
    theme_minimal() +
    theme(
      legend.title = element_blank(), 
      legend.text = element_text(size = 10),
      plot.background = element_rect(fill = "white", color = NA),  # Ensure white background
      panel.background = element_rect(fill = "white", color = NA)  # Ensure white background
    )
  
  # view and save replicates as .png
  print(p)
  ggsave(paste0("plot_", gsub(" ", "_", rep), ".png"), plot = p, width = 8, height = 6, bg = "white")
}
